---
output: 
  github_document
--- 

# Saddlepoint Approximation

Zachary R. McCaw <br>
Updated: 2023-06-26

```{r global_options, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
```

## Description

This package performs the linear model score test with p-value calculation via the empirical saddlepoint approximation. 

## Installation

```{r, eval=FALSE}
devtools::install_github(repo = "zrmacc/Saddlepoint")
```

## Usage

### Example data

Simulate example data for `n = 1e3` subjects where the proportion of phenotypic variation explained by genotype is 0% and the proportion by the covariate is 20%.

```{r, cache=FALSE}
set.seed(100)
data <- Saddlepoint::DGP(n = 1e3, pve_g = 0.0, pve_x = 0.2)
```

### Fit the null model

Estimate the null model which excludes genotype. Extract the mean `mu` and residual variance `sigma2` under the null.

```{r, cache=FALSE}
fit <- stats::lm(y ~ x, data = data)
mu <- fit$fitted.values
sigma2 <- stats::sigma(fit)^2
```

### Perform score testing

* The required arguments are genotype `g`, phenotype `y`, and the null model mean `mu` and residual variance `sigma2`. 
* The covariates `x` should either be supplied directly to the score test (as below), or else regressed from `g` prior to performing the score test. 
* The output includes a flag `used_sp` of whether saddlepoint approximation was applied. If `FALSE`, the p-value was calculated via normal approximation. By default, normal approximation is applied when the magnitude of the score statistic is less than 1.96 (the 97.5th percentile of the normal distribution).

```{r, cache=FALSE}
result <- Saddlepoint::ScoreTest(
  g = data$g,
  mu = mu,
  sigma2 = sigma2,
  x = data$x,
  y = data$y
)
show(result)
```
